\documentclass{article} 
%\documentclass[draft]{article}  % See Skeleton.tex for examples of many things

\usepackage{underscore} % accepts  _ in text mode
% Comment the next two so can have explicit new commands for html
% \usepackage{ifpdf} % detects if processing is by pdflatex
% \usepackage{/home/hkieffer/xtex/newcom}  % Hughs conventions

% local definitions
\textheight=9.6in  \topmargin=-0.5in 
\textwidth=7.70in  \oddsidemargin=-0.7in \evensidemargin=-0.7in
\parindent=0.em \parskip=1.ex %  no indent & paragraph spacing

%\newcommand{\short}{full}    % dummy for format
\newcommand{\qi}{\\ \hspace*{2.em}}      % indent 1
\newcommand{\qii}{\\ \hspace*{4.em}}     % indent 2
\newcommand{\qiii}{\\ \hspace*{6.em}}    % indent 3

\newcommand{\np}{\textbf}  % name of program or routine
\newcommand{\nf}{\textit}  % name of file
\newcommand{\nv}{\texttt}  % name of code variable in text
\newcommand{\nvf}{\mathtt} % name of code variable in equation
\newcommand{\nj}{\textsf}  % name of input parameter in text
\newcommand{\njf}{\mathsf} % name of input parameter in equation
\newcommand{\qd}{$^\circ$}        % degree symbol
\newcommand{\qb}{\\ $ \bullet \ $}        % bullet subsection

\title{KRC planetary surface temperatures: Helplist}
\author{Hugh H. Kieffer \ \ File=-/krc/Doc/helplist.tex Version 3.6.5 2020apr13}
%_History
% 2016feb08  HK Update spare parameters
% 2016nov21  HK 3.4.3
% 2018jan30  HK v 3.5.5
\begin{document}
\maketitle

\tableofcontents
%\listoffigures
 \listoftables
% \pagebreak

\section{Introduction} %------------------------------------------------

This document is intended to help the expert user of KRC set up an input file
that addresses their goals and will generate the kind of output they desire. It
is assumed that the user is familiar with the KRC journal article: H.H. Kieffer,
Thermal model for analysis of Mars infrared mapping, J. Geophys. Res.: Planets,
v.118, 451-570 (2013) [Ref. 1]

The evolution of KRC code is contained in:  \nf{evolve.txt}

A crude diagram of the call architecture is in:  \nf{flow.txt}

Capabilites new in version 3.4.x are decribed in: \nf{V34UG.tex} (or .dvi or .pdf versions)

Eclipse capabilites (V3.5) are described in: \nf{eclipse.tex}  (``)

[[ Development for thermal beaming of asteroids is described in: \nf{Beaming.tex} (``) ]] not released

\subsection{Notation use here}
The following fonts styles have been partially implemented: 
\qi File names are shown as \nf{file}. 
\qi Program and routine names are shown as \np{PROGRM [,N]} 
\qii where \np{N} indicates a major control index. 
\qi Code variable names are shown as \nv{variab} and within equations as $\nvf{variab}$.  
\qi Input parameters are shown as \nj{INPUT} and within equations as $\njf{INPUT}$

When KRC was made double-precision, an '8' was added to many routine names, both
as the source-code file and as they are called; this 8 is commonly omitted in
this documentation.

\section{METHOD}

The program is designed to compute surface and subsurface temperatures for a
global set of latitudes at a full set of seasons, with enough depth to capture
the annual thermal wave, and to compute seasonal condensation mass. For historic
reasons, the code has substantial optimization. Although developed for Mars,
there are generalities that allow this code set to be used for any solid body
with any spin vector, in any orbit (around any star), with or without an
atmosphere; this is also the source of some of the complexity.
 
Method is explicit forward finite differences with exponentially increasing
layer thickness and binary time increase with depth where allowed by stability.
Depth parameter is scaled to the diurnal thermal skin depth.  Initially starts
at 18 hours with the mean temperature of a perfect conductor.  Second degree
perturbation is applied at the end (midnight) of the (third) day; this jumps the
mean temperature of all layers and the lower boundary to equal the mean surface
temperature.

Boundary condition treatment:
\qi  Perturbation solution of quartic equation at surface for each iteration;
    temperature gradient assumed uniform in top interval.
\qi  Lower boundary may be insulating or constant-temperature, or fixed heatflow

Atmospheric Radiation: KRC uses a one-layer atmosphere that is grey in both the
solar and infra-red regions. The default atmospheric
parameters are based on estimates of Mars' gas and aerosol properties.
\qi  Two-stream Delta-Eddington model for insolation; direct onto sloped surface 
and diffuse, with possible twilight extension.
\\ The atmosphere temperature is based on Delta-Eddington solar absorption and IR opacity

Keplerian orbital motion; seasons are at uniform increments of time. Mean 
orbital elements are pre-calculated for any epoch (all planets and several
comets and asteroids) by the PORB code set.

Units are SI, except for use of days for orbital motion and rotation period 

Note: Insolation at a season is computed on the assumption that the body is
spinning at a fixed orbital position for that season. This is equivalent to
assuming the ``day'' (sol) is short compared to the ``year'' (orbital
period). This is equivalent to all longitudes having the same insolation. Hence,
KRC is not appropriate for bodies with small synchronous rotations to their
year, such as Mercury.

Options:
\qi  Different Physical properties below a set layer (IC). 
\qii May define many zones via an input table
\qi  Regional slope. Self- or far-field heating
\qi  Three (for Mars) ways to handle seasonal global pressure variation .
\qi  Up to two zones of temperature-dependent conductivity and specific heat
\qi Non-Lambertion solar absorption.
\qi Global geothermal heat flow.
\qi Eclipses by the planet or satellite.
\qi Incident flux from a planet onto a satellite, thermal is sinusoidal.
\qi For sloped surfaces, the ``far-field'' radiation can be defined by a prior KRC run.

Atmosphere condensation: 
\qi Global integral of frost-gas budget can control surface pressure.
\qii Can have virtually any single condensing gas desired.
\qi Allows different surface elevation for each latitude zone.
\qii  Zonal frost saturation temperature tracks local surface pressure.
\qi Option for cap albedo to depend upon mean daily insolation.
\qii Frost albedo can vary continuously with frost amount

\subsection{Convergence Notes} %-----			

Convergence prediction routine can't jump more than one time constant
(TAU=X**2/2) \ $\tau =x^2/2$ for the total thickness.  Therefore, if X(N1) is
small, make DDT smaller than usual.  If DELJUL is much smaller than (X(N1))**2/2
\ $X_{N1}^2/2$, then DDT can be as large as 0.3.  Otherwise DDT must be about 0
for the prediction routine to work well (it assumes the 3rd derivative to be 0).

\section{INPUT FILE}

KRC asks for the name of an input file, default is \nf{krc.inp}, and an output
file, default is \nf{krc.prt}. If the desired file is not in the current
directory, then its name must be surrounded by single quotes.

All parameters for KRC are set by a formatted text file.  An example is
master.inp , which has default values for a 19 latitude set for a run of three
martian years, with the last output to disk. Parameter values are listed below
their titles, which are in many cases identical to the code name, and last
character of the title is above the last location in the field. Thus, integer
values MUST be aligned. Titles with a leading "[" indicate that the value is not
  used. The recommended procedure is to copy master.inp and edit only the values
  you wish to change. The number of lines of Latitudes and Elevations must match
  the value of N4, e.g., 2 lines for N4=11:20, entries beyond the N4 position
  may be left blank or contain the end of the line. The 7 lines following
  Elevations are a geometry matrix, default is for Mars orientation and orbit in
  2010. This matrix should not be edited; but can be replaced by the matirx for
  other objects by the results of running PORBMN .

The parameter title lines are skipped, so that you could put comments there
carefully.

The first input line is always KOLD,KEEP [I*], which sets file usage; these are
described in \S \ref{dbf}. If KOLD=0, then a full set of input values is read.
\qi If and only if there is a third non-zero integer, then KRC will read the next
card as six debug flags, IDB1 to IBD6, which are normally zero.
  See \S \ref{debug} below \\
The next (normally second) line is free text where you can outline the purpose 
of your run; up to 84 characters.

Change lines may follow immediately after the geometry matrix (see \S \ref{pc})
. The end of definition of a "case" is indicated by a "0/" line. Two successive
"0/" lines ends the run.

The source code for 'krcc8m.f' indicates for many of the parameters which
subroutine sets it; as the routine name in lowercase just below the parameter
name.
\vspace{-3.mm} 
\begin{verbatim} 
 - - - - - - - - - - - - - - - - - - - -
Type 1	Real parameters  (8F10.2) ================================
  Surface Properties
1   ALB    Surface albedo
2   EMIS   Surface emissivity
3   SKRC   Surface thermal inertia [J m^-2 s^-1/2 K^-1] { cal cm * 4.184e4}
4   COND2  Lower material conductivity (IC>0)
5   DENS2  Lower material density (IC>0)
6   PERIOD Length of solar day in days (of 86400 seconds)
7   SPHT   Surface specific heat [J/(kg K)]  {cal/(g K) * 4184.}
8   DENS   Surface density [kg/m^3] {g/cubic cm. *1000}
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
  Atmospheric Properties
9   CABR    IR opacity of dust-free atmosphere of PTOTAL surface pressure
10  AMW     Molecular weight of the atmosphere
11  ABRPHA  First Clausius-Clapeyron coefficient  SatPrA 
12  PTOTAL  Global annual mean surface pressure at 0 elev., Pascal[=.01mb]
             If <=1.; no atmosphere at all.
             If KPREF=2, global average of atmosphere plus cap system.
13  FANON   Mass-fraction of mean atmosphere that is non-condensing
14  TATM    Atm temp for scale-height calculations
-   -   -   -   -   -   -   -   -   -   -   -   -   -   -   -   -   -   -   
15  TDEEP   Fixed bottom temperature. Used if IIB>=1.
16  SPHT2   Lower material specific heat (IC2>0)
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
  Dust  & Slope Properties
17  TAUD    Mean visible opacity of dust, solar wavelengths
18  DUSTA   Single scattering albedo of dust
19  TAURAT  Ratio of thermal to visible opacity of dust
20  TWILI   Twilight extension angle [deg]
21  ARC2    Henyey-Greenstein scattering asymmetry factor
22  ARC3    Minimum allowable numerical convergence factor   Theory is 1.
23  SLOPE   Ground slope, degrees dip. Only pit may slope beyond pole.
24  SLOAZI  Slope azimuth, degrees east from north. <-360 is a pit
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
  Frost Properties
25  TFROST  Minimum Frost saturation temperature
            may be overridden by local saturation temperature (LVFT)
26  CFROST  Frost latent heat [J/kg] {cal/gm*4184. [ Not used if
27  AFROST  Frost albedo, may be overridden (LVFA) [ TFROST never
28  FEMIS   Frost emissivity                       [ reached
29  AF1     constant term in linear relation of albedo to solar flux
30  AF2     linear term in relation of albedo to solar flux units=1/flux
              Afrost = AF1 + AF2 * <cos incidence> SOLCON / DAU^2
31  FROEXT  Frost required for unity scattering attenuation coeff. [kg/m^2]
            the greater of this and 0.01 is always used.
32  fd32    Second Clausius-Clapeyron coefficient   SatPrB
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
  Thermal Solution Parameters
33  RLAY    Layer thickness ratio
34  FLAY    First physical layer thickness (in skin depths)
35  CONVF   Safety factor for classical numerical convergence
              0 for no binary time division of lower layers
             >0.8 for binary time division. Larger is more conservative
36  DEPTH   Total model depth (scaled) (overrides FLAY if not 0.)
37  DRSET   Perturbation factor in jump convergence. If = 0., then
              all layers reset to same average as surface layer. Else,
              does quadratic curve between surface and bottom averages
38  PHOG     Reflectance photometric function control
               0=Lambert  <-1=Lommel-Seeliger  +=Kheim-Vasavada
               else (-1< x <0) Minnaert with exponent -x (is positive)
               See V34UG.pdf, Sections 5 and 13
39  GGT     Surface boundary condition iteration test on temperature
40  DTMAX   Convergence test: RMS layer T changes in a day
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
  Orbit Geometry & Constants
41  DJUL    Starting Julian date of run -2451545 (J2000.0) (N5>0)
             If LKEY (logical item 12) is set True, this is desired Ls of 
             first disk output date (set by JDISK: integer item 12)
42  DELJUL  Increment between seasons in Julian days (if N5>1)
43  SDEC    Solar declination in degrees. (if Not LPROB)
44  DAU     Distance from Sun in astronomical units (if Not LPROB)
45  SUBS    Aerocentric longitude of Sun, in degrees. For printout 
             only. Computed from date unless N5=0 (for printout only)
46  SOLCON  Solar constant. Total stellar irradiance at 1 AU  [W/m^2]
47  GRAV    Surface gravity.  MKS-units
48  AtmCp   Specific heat at constant pressure of the atmosphere [J/kg/K]
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
  Temperature dependent conductivity.  Ignored unless LKOFT set.
49  ConUp0  Constant coef for upper material 
50  ConUp1  Linear   in k=c0+c1x+c2x^2+c3x^3 where x=(T-220)*0.01
51  ConUp2  Quadratic    " 
52  ConUp3  Cubic coeff. "
53  ConLo0  Constant coef for lower material 
54  ConLo1  Linear      as for ConUp above
55  ConLo2  Quadratic    "
56  ConLo3  Cubic coeff. "
  Temperature dependent specific heat.  Ignored unless LKOFT set.
57  SphUp0  Constant coef for upper material 
58  SphUp1  Linear   in k=c0+c1x+c2x^2+c3x^3 where x=(T-220)*0.01
59  SphUp2  Quadratic    " 
60  SphUp3  Cubic coeff. "
61  SphLo0  Constant coef for lower material 
62  SphLo1  Linear      as for   SphUp above
63  SphLo2  Quadratic    "
64  SphLo3  Cubic coeff. "
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
  Computed REAL values
65  HUGE    = 1.D308    large  REAL*8 constant with margin
66  TINY    = 1.D-307   small  REAL*8 constant with margin
67  EXPMIN  = 700.d0    safe negative R*8 exponent
68  YRIDAY  Length of a year in days
69  FLOST   Atm frost 'lost' in the atm. in last day at current lat./season
            v356 and later, this is spare.
70  RGAS    = 8.3145  ideal gas constant  (MKS=J/mol/K)
71  TATMIN  Atmosphere saturation temperature
72  PRES    Local surface pressure at current season
73  OPACITY Solar opacity for current elevation and season
74  TAUIR   current thermal opacity at the zenith
75  TAUEFF  effective current thermal opacity 
76  TATMJ   One-layer atmosphere temperature
77  SKYFAC  fraction of upper hemisphere that is sky
78  TFNOW   frost condensation temperature at current latitude
79  AFNOW   frost albedo  at current latitude
80  PZREF   Current surface pressure at 0 elevation, [Pascal]
81  SUMF    Global average columnar mass of frost [MKS]
82  TEQUIL  Equilibrium temperature (no diurnal variation)
83  TBLOW   Numerical limit (Blowup) temperature
84  HOURO   Output Hour requested for "one-point" model
85  SCALEH  Atmospheric scale height
86  BETA    Atmospheric IR absorption
87  DJU5    Current Julian date (offset J2000.0 ala PORB convention)
88  DAM     Half length of daylight in degrees
89  EFROST  Frost on the ground at current latitude [kg/m^2] {g/cm^2 * 10.} 
90  DLAT    Current latitude
91  COND    Top material Thermal conductivity (for printout only)
92  DIFFU   Top material Thermal diffusivity (for printout only)
93  SCALE   Top material Diurnal skin depth (for printout only)
94  PIVAL   pi
95  SIGSB   Stephan-Boltzman constant (set in KRC)
96  RADC    Degrees/radian

Type 2 Integer Parameters (8I10) ====================================
 To simplify coding logic, some may be input as 0, meaning none or never,
     but will be modifed automatically to very large, Indicate by: -->never 

1   N1      # layers (including fake first layer) (lim MAXN1)
2   N2      # 'times' per day (lim MAXN2). Must be an even number, 
             should be a multiple of N24 and NMHA.
3   N3      Maximum # days to iterate for solution (lim MAXN3)
             This can be 1, but then must use  DELJUL ~= PERIOD
             If N3 lt 3, first day starts on midnight. else at 18H 
4   N4      # latitudes (lim MAXN4). Global integrations done for N4>8
5   N5      # 'seasons' total for this run. If 0, then DAU and SDEC will be 
             used as entered for a single season.
6   N24     # 'hours' per day stored, should be divisior of  N2 (lim MAXNH)
7   IIB     Bottom control: 0=insulating, -1=constant temperature 
             -2=start all layers =TDEEP & constant temperature 
              += geothermal heat flow in milliwats/m^2
8   IC2     First layer (remember that 1 is virtual) of changed properties. 
             <3 will be changed to 999 =homogeneous
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
9   NRSET   # days before reset of lower layers first season;  >N3=no reset
10  NMHA    # 'hour angles' per day for printout (no limit)
11  NRUN    Run #; appears in some printout. Initalized as 0 and   
             auto-increment whenever disk file opened. May be modified
12  JDISK   Season count that disk output is to begin. <1 -->never
13  IDOWN   Season at which to read change cards
14  I14     Index in FD of flexible print
15  I15      "" and >100 is code for special output to fort.77 via tun.f
16  KPREF   Mean global pressure control. 0=constant
              1= follows Viking Lander curve  2=reduced by global frost, but
              then N4 must be >8, and latitudes must be monotonic increasing
              and must include both polar regions (no warning for your failure)
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
17  K4OUT   Disk output control: See details in DISK BINARY FILES section
             Three modes of direct access Fortran files;  one case per file.
               -=KRCCOM(once), then TSF [,TPF [,TAF]]
                 -1 saves TST, the surface kinetic temperature
                 -2 also saves TPF, the top-of-atmosphere bolometric temperature
                 -3 also saves TAF, the 1-layer atmosphere kinetic temperature
               0=KRCCOM,LATCOM each season
               1:49=KRCCOM,DAYCOM for the last latitude; each season
              Modes of bin5 file for multiple cases. ONLY 52 IS STILL SUPPORTED
               x51=(Hours, 2 min/max,  lat, seasons, cases)
               52=(hours,  7 items, lat, seasons, cases) 
               x54=(many seasons, 5 items,lats, cases)
               x55=(many seasons,9 items, cases)
               x56=(packed T hour and depth, latitude,season,case)
18  JBARE   J5 season count at end of which to set frost amount to 0. 0-->never
19  NMOD    Spacing of season for notification. minimum of 1
20  IDISK2  Last season to disk for which TDISK prints notice
             Note: Special routines MKRC,KRCA and TYEARP use TDISK differently
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Computed Integer values   But note 24 may be input
21  KOLD    Season index for reading starting conditions
22  KVALB   Flag: to use seasonal surface albedo ALB
23  KVTAU   Flag: 1:TAUD=SEASTAU(SUBS)  2:CLIMTAU opacities for dust and ice
24  ID24    Flag: Direct Access file word type: 4=R*4 else=R*8
     this single item in 21:40 can be set by a change line: 2 24 x 'ID24' /
25  NUMVER  Version number as an integer
26  KFARAC  Sets reporting of TFAR8 reads
27  NBKRC   Number of bytes in KRCCOM,   not used anywhere
28  NFD     Number of real items read in
39  NID     Number of integer items read in
30  NLD     Number of logical items read in
31  N1M1    Temperature vrs depth printout limit (N1-1)
32  NLW     Temperature vrs depth printout increment
33  JJO     Index of starting time of first day
34  KKK     Total # separately timed layers
35  N1PIB   N1+IB Used to control reset of lowest layer
36  NCASE   Count of input parameter sets in one run
37  J2      Index of current time of day
38  J3      Index of current day of iteration
39  J4      Index of current latitude
40  J5      Index of current "season"

Type 3 Logical Parameters (10L7) ====================================

1   LP1     Print program description. TPRINT(1) 
2   LP2     Print all parameters and change cards (2) for current case
3   LP3     Print hourly conditions on last day (3), every lat, every season
4   LP4     Print daily convergence summary (4)
5   LP5     Print latitude summary (5)
6   LP6     Print TMIN and TMAX versus latitude and layer (6)
7   LPGLOB  Print global parameters each season
8   LVFA    Use variable frost albedo. Uses AF1 & AF2 (real # 29,30)
9   LVFT    Use variable frost temperatures
10  LKOFT   Use temperature-dependent conductivity and specific heat
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
11  LPORB   Call PORB1 just after full input set
12  LKEY    Treat DJUL (real item 41) as the Ls of the first output
             season, set by JDISK [the 12'th integer input]
13  LSC     Read change cards from input file at start of each season
14  LZONE   Use a zone depth table file to set layers.
15  LOCAL   Use each layer for scaling depth
16  LD16    Print hourly table to fort.76 [TLATS] 
              SUBS,DLAT,ALB,SKRC,TAUD,PRES at end of every latitude
              and every hour:  Hour  T_K  IR_Inc  Solar  T_Planet
17  LD17    Print temperatures at end of last day of each season
18  LD18    Print of parameters and layers was done after last change
19  LD19    Output to fort.79 [TLATS] insolation and atm.rad. arrays 
            for every time-step for every latitude, every time tlats is called.
20  LONE    (Computed) Set TRUE if KRC is in the "one-point" mode
================================================================

followed in 'krccom' by: 
84 character KITLE	Description/purpose of this run.
20 character DAYTIM	run  date and time
================================================================

Latitude(s) (10F7.2)   N4 latitudes in degrees
Latitudes to be in order; south to north. [[If last latitude is
.LE. 0, will assume symmetric results for global integrations]]

Elevation(s) (10F7.2)  N4 values in Km corresponding to latitudes

Orbital Parameters (LPORB=T) Format identical to that produced by PORB
program set ASCII file output. So these can be directly pasted with an
editor. See porbc8m.f
 - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
\end{verbatim}

\section{PARAMETER CHANGES \label{pc}}

Fortran list-directed.  Change the values in KRCCOM.  Lines of four (or more) 
white-separated values, a "/" terminates the read and leaves remaining values
unchanged.  The 4 required items are:
\vspace{-3.mm} 
\begin{verbatim}
    1: Type (integer): see table below
    2: Index in array (integer): as listed in Input File table above
    3: New value, numeric:  0.=false. Will read as real and convert as needed.
    4: File name or a reason:  Text string within single quotes; or a /
         Missing quote may cause run failure.
   [after a / (forward slash) nothing is read, so you can use for comments]
\end{verbatim}
The print file will list each change as read, followed by the title of the
changed item. It is a good idea to look at this print to be sure you changed
what you intended.

\begin{verbatim}
 Type     Meaning                                      Valid Index

   0   End of Current Changes                              any
   1   Real Parameter                                     1:NFDR=64
   2   Integer Parameter                                  1:NIDR=20
   3   Logical Parameter                                  1:NLDR=20
   4   New Latitude Card(s) Follow                         any
   5   New Elevation Card(s) Follow                        any
   6   New Orbital Parm Cards Follow (LPORB Must be True)  any
   7   Text becomes new Title                              any
   8   Text becomes new output or input file name. Value ignored if not mentioned
         if index=3, name of far-field temperature file to read
                       value 0/1 sets report of each read off/on
         if index=4, name of far-field temperature file to read for held Tatm.
         if index=5, name of type 5x (52) packed output file
         if index=21, name of output direct-access file. Lasts for one case
         if index=22, call SEASALB to read variable ALBEDO
         if index=23, call SEASTAU to read variable TAUD
         if index=24, call CLIMTAU to read Mars climate
         if index=25, name of new zone table to read
   9   Complete new set of input follows                   any
  10   Text becomes new One-Point input file name
  11   This is a set of parameters for "one-point" model 
          For this type, 9 values must appear in a rigid format
  12   Set of 2*4 coefficents for T-dep. conductivity.  List-directed IO
  13   Set of 2*4 coefficents for T-dep. specific heat. List-directed IO 
  14   Set of 12 values for eclipse. See -/Doc/v35/eclipse.pdf sec. 5.5 
  e.g. 14  1 5.2026 71492. 0.6711D6 3121.6 3.551 0.01 6000. 12. 2 77 77 / Europa
  15   Set of 7 values for flux load from primary body (planet). " sec. 7.1
  e.g. 15  0.62 0. 0.  0.21 0.21  0. 12. / Jupiter flux on Europa, nearside center
  16   Latitude and name for high-time-resolution surface temperature file
         See eclipse.pdf section 8
\end{verbatim}
 
\subsubsection{Type 8 notes}

Each of the index actions can be terminated by using a file name of less than 4
characters; .i.e. ``off''.

For Type 8, SEASALB and SEASTAU read 2-column, white-separated text files. \\
To start variable albedo, use input card: 
\qi  8 22 0 'AlbedoFileName' / Variable albedo text file name \\
Can revert to constant albedo by using a non-existant file name. E.g.,
\qi  8 22 0 'badName' / turn variable albedo off \\
Text table files of value versus season will be read at the start of a
run. These will apply to ALL latitudes. See example  \nf{valb1.tab}   \\
Variable dust opacity done the same way, with 22 being replaced with 23 \\

CLIMTAU files have dust and ice opacity over season and latitude. Uses BINF5 to
read a binary array (72 seasons, 36 latitudes, 2=dust/ice) of opacities. The
sample file \nf{THEMIS1yearDustIce.bin5} is described in section [159] of Ref 1.

Index=3 or =4 will read type -n file to be used for far-field surface and atmosphere.  See \S \ref{fff}

Index=5 writes a dense type 52 file that can hold many cases and contains virtually all the input and output values! See \S \ref{type52}. It cannot be input to KRC but an IDL reader \np{readkrc52} is provided.

\subsubsection{Other types}
For Type 12 and 13, 8 white-space-separated coefficients must follow after 
the type on the same line, with no intervening index or text 

Type 14, 15, 16 apply to all following cases until a corresponding change line
with the 2nd item 0 .
 

\section{Contents of COMMOMS } %_______________________
 COMMON /KRCCOM/ \ Input and transfer variables. See krcc8m.f  
\\ COMMON /DAYCOM/ \ Layer and time-of-day items. See dayc8m.f  
\\ COMMON /FILCOM/ \ File names. See filc8m.f  
\\ COMMON /HATCOM/ \ Store post-2003 items. See hatc8m.f  
\\ COMMON /LATCOM/ \ Latitude-dependent items. See latc8m.f  
\\ COMMON /PORBCM/ \ PORB system geometry matrix.  See porbc8m.f  
\\ COMMON /UNITS/ \ Logical units for I/O and errors.  See unic8m.f  \\

Because the binding routines to IDL are intolerant of any errors, changes to
KRCCOM, DAYCOM and LATCOM are avoided if possible. Rather, in 2004July HATCOM
was added as a "catch-all" for any new items.

A listing of all Fortran commons can be generated by these Linux commands: \\ 
cd /home/hkieffer/krc/src [replace top part of path with local installation] \\ 
rm allinc.txt \\ 
cat  *c8m.f $>$ allinc.txt 
\pagebreak
              
\section{Error Returns} %_______________________

If you run into errors that are not related to blunders in the input file,
please contact your source for KRC.

\subsection{``Case had fatal error''} %-----------------------
Message to both the print and error files, the value printed is explained in Table \ref{ercode.tab} 

\begin{table} 
\caption[Return codes]{Return codes, as seen at KRC main, after return from TSEAS, and maybe from lower routine:}
\label{ercode.tab}
\rule[.05in]{7.in}{.01in}
\begin{tabbing} 
WWWW \= WWW \= WWWWWWWW \=   \kill 
in  \>  value\> Comment  \\
TCARD \> \>  2:19 \\
   \>  2 \> restarted from disk record \\
   \>  3 \> continue from current conditions \\
   \>  4 \> Switch to "one-point" mode \\
   \>  15 \> END of data  \\
   \>  16 \> Error reading buffer of one-point or T-dependent coefficients \\
   \>  17 \> Early End Of File for above \\
TDAY8(1 \> \> initiation  [by +20] \\
   \>   23 \> Some layer unstable   \\
   \>   24 \> Too many Layers generated by zone table \\
   \>   27 \> TFINE initiation failed  \\
TLATS \> \> 21:29 from lower TDAY(2 [by +20] \\
   \>  22 \> numerical blowup \\
   \>  25 \> ECLIPSE failed \\
   \>  28 \> ECLIPSE reaches midnight (possible near pole of body) \\
   \>  31:36 \> Error n-30 in TFAR, see below [by +30] \\
   \>  37 \> Using far-field file, but N2 is not an integral multiple of the number of hours in the fff. \\
   \>  38 \> Using fff, but that has no match to current latitude \\
   \>  39 \> ECLIPSE return error at current latitude. \\
TSEAS \>  \> none in TSEAS itself \\
TFAR \> \> Error when arg3 (R*8) returns negative: use magnitude +40 \\
  \> 40+ \>  magnitude below, +3 for 2nd file \\
   \> 1  \>   request for file information, but none open \\
   \> 2  \>   invalid number of temperature types (records per season) \\
   \> 3  \>   an I/O error occured in TFAREAD \\ 
TDISK \> \> In TDISK, will STOP. Error opening Direct Access file. See also \S \ref{edisk} \\
\end{tabbing} 
\rule[.25in]{7.in}{.01in}
\vspace{-.1in}
\end{table}

\subsection{Tday Blowup} %------------------------------------
In initiation step TDAY8(1

 (QQ.LT. ARC3): convergence factor less than safety factor specified by input
iten 1:22, which should never be less than 0.8. In theory, this factor should be
1.0, but KRC is generally stable down to 0.8 for temperature-independent
materials. For temperature-dependent materials, KRC uses the values at the
forecast global mean temperature; if this error occurs, increase ARC3 and the
layer thicknesses.

 (TSUR.LE. 0. .OR. TSUR.GT.TBLOW): Surface kinetic temperature outside
reasonable range. \nv{TBLOW} is set (in TLATS) as twice the equilibrium
temperature for the surface perpendicular to the specified insolation at the
current season.

In either case, control goes to 340 in TDAY; iteration is terminated and several
values are sent to the print file. TDISK is called to output the current season
and to close the file. TPRINT is called to print out the full input set and the
latest daily convergence. The return code is set to 2.

\subsection{TDISK error explanations \label{edisk} } %--------------------
Below are extracts from the FORTRAN code. Material between single quotes is printed; items following on the same line are variables whose value will be printed. 
\qi 'P' goes to print file. \ \ 'E' goes to error log.
\qi 'DA'  is direct-access \ \  @ indicates action taken
\\ 
\\ P: 'TDISK Error:  KODE,KREC,JREC=',KODE,KREC,JREC
\\ E:  ' TDISK:1, ERROR OPENING FILE. IOSTAT=',IOS
\\ E:   '  IOD2=',IOD2,'  status=',CSTAT,'  recl=',NRECL
\\ E:   '  file=  ',FDIRA
\qii  Trying to open DA file, but failed: reading a file that does not exist or writting a file that already exists.
\qii @ KRC stops
\\ 
\\ E: 'TDISK: K4OUT invalid, will use -3:',K4OUT
\qi Trying to open a DA file, but the value of K4OUT in the input file invalid.
\qii  @ Locally sets type to .tm1 and proceeds
\\ 
\\ E: 'TDISK: Record smaller than KRCCOM. R*8 words:',NWTOT,NWKRC
\qi Some change to the source code krcc8m.f has generated a file record length smaller than KRCCOM. 
\qii Fatal, @ will stop
\\ 
\\ P:  'Bad size change, CLOSE file. See Error File'
\\ P: 'Closed direct access file. Type=',ITOP
\\ E: 'TDISK: Invalid Change of N24, NLAT, or \#to Disk for K4OUT=52'
\\ E: 'TDISKe:',N24,MM1, N4,JJJ(4),J,JJJ(5)
\qi Change to any of these would change the size of a Type 52 array, not allowed.
\qii @  Closes the file, file will be unreliable
\\ 
\\ E:  ' TDISK:2, WRITE, BUT NO FILE OPEN'
\qi   Trying to write a seasonal record, but file is not open
\qii @ Will continue calculations
\\ 
\\ E: 'TDISK:3, READ, BUT NO FILE OPEN'
\qi Trying to read a DA record, but the file-open earlier failed 
\qii @ Will continue calculations, results unreliable  
\\ 
\\ E: 'TDISK:4, no file was open'
\qi Asked to close a file that was not open.   No action.
\\ 
\\ E: 'TDISK:5,wrong conditions: LOPN2,ITOP',LOPN2,ITOP
\qi  Trying to write KRCCOM to a DA file, and either the file was not open or the open file was the wrong type. No other action.
\\   
\\ E: 'TDISK:7, Asked to close but LOPN4 false',JREC
\qi Some routine wants to write a Type 52 file, but contents are not defined.
\qi Should never occur; programming error. @ KRC continues
\\ 
\\ E: 'TDISK:5, Called with invalid KODE=',kode
\qi Should never occur; programming error. @ KRC continues


\section{DISK BINARY FILES \label{dbf} } %_______________________

The routine TDISK is used to read or write seasonal/global direct-access binary
files and to write multi-case type 52 (.bin5) files. The first season to write
is specified by JDISK, all following seasons will go to the same file. For
direct-access files type 0 or positive, each file record consists of KRCCOM plus
LATCOM or KRCCOM.  plus DAYCOM. Type -n have one record the contines KRCCOM
followed by n records of temperature for each season. Writing of type 52 and
direct-access files are independent of each other and may be simultaneous.

Binary files are REAL*8, with the single exception that type -n may be written
as REAL*4 (to save space for really large runs) by setting ID24 to 4 (see \S
\ref{pc}; R*4 files can be read with the TFAR and TFAREAD routines, which will
auto-detect the precision; no other use of binary R*4 files with KRC versions
3.1.1 or later has been tested!

\subsection{Items which control file I/O } %------------
\vspace{-3.mm} 
\begin{verbatim}
KOLD & KEEP on first input line
  KOLD: 0= input card set follows;  else=disk record number to start from,
         then will read any change cards.
        If LPORB in old file was True, then there must be a PORB card set 
          as the set of lines following the KEEP,KOLD line
  KEEP: 0= close disk file after reading seasonal record KOLD;
       >0= value of JJJJJ at which to start saving seasons in same disk 
        file [overrides JDISK].
  To start from a prior seasonal run, need to determine the record 
  corresponding to the desired season;
         KOLD=J5_target - JDISK(old) ; >0
         set KEEP=1, change card J5=number of new seasons, set K4OUT.

JDISK sets the first season to save results

N5    sets the last season to run

K4OUT sets the direct-access record content: (ignored for Type 52)
 -1:-3  First record is COMKRC=/KRCCOM/= FD(96R8),ALAT(37R8),ELEV(37R8)
         ,ID(40I4), LD(20L4), KITLE(C84), DAYTIM(C20)   Sizes set in krcc8m.f
         Then records of TSF, [TPF, [Tatm]]
          For each season. Each record is MAXNH,MAXN4 = 96*37 = 3552 real*8 words
 0      Will output records of KRCCOM+LATCOM for each season
 1:49   Will output records of KRCCOM+DAYCOM for the last computed latitude.
\end{verbatim}

If a change ' 8 5 0 $<$filename$>$ ' has been set, then will write a type 52
bin5 file at the end of a run, with dimensionality of 5. This allows multiple
cases, each with a "prefix" for each case consisting of 4 size integers
(converted to Float) followed by KRCCOM; after this may come vectors of
parameters versus season. The next-to-last dimension is increased to allow room
for the prefix to be embedded in the bin5 array.  Each dimension is adjusted to
the necessary size. Each case has the same structure; this simplifies coding
although some items are then present redundantly.

KRC input items that would increase any of the bin5 dimensions are not allowed
to change between cases in a file. Decrease in these sizes are allowed, however
this will leave regions of the file undefined and the only clue will be the
values in krccom that are stored for each case. Increases to larger than the
value for the first case will cause an error and closing of the file. This
restriction is on:
\qi N5-JDISK: number of output seasons
\qi N4: number of latitudes
\qi N24: number of hours output

Although KRC allows the N1, the number of layers, to change between cases the
IDL type 52 reader \np{readkrc52} will only extract the number in the first
case, so N1 should not increase between cases.

The first 4 words of the prefix, and thus of the bin5 array, are:
\vspace{-3.mm} 
\begin{verbatim}
(1)=FLOAT(NWKRC)   ! Number of words in KRCCOM
(2)=FLOAT(IDX)     ! 1-based index of dimension with extra values
(3)=FLOAT(NDX)     ! Number of those extra
(4)=FLOAT(NSOUT)   ! [Available of other use]
\end{verbatim}

See also Appendix \ref{type52} for a detailed description and how it is unpacked.

\vspace{-3.mm}
\begin{verbatim}
    52=(N24 hours, 7 items, N4 lats, NDX+ seasons, cases)
The prefix section contains: sub_array(seasons,5)(0-based index)
  0)=DJU5  1)=SUBS  2)=PZREF  3)=TAUD  4)=SUMF

The 7 items are:  1)=TSF  2)=TPF  3)=TAF  4)=DOWNVIS  5)=DOWNIR
6) packed with [NDJ4,DTM4,TTA4,      followed by TIN(2+
7) packed with [FROST4,AFRO4,HEATMM, followed by TAX(2+
  The number of layers for TIN and TAX is the smaller of: the number computed 
and that fit here.

---------- The other type 5x listed below are no longer supported ------

    51=(N24 hours, 2: TSF TPF, N4 lats, NDX+ seasons, cases)
The prefix is identical to Type 52
  0)=DJU5  1)=SUBS  2)=PZREF  3)=TAUD  4)=SUMF

    54= (seasons, 5 items, NDX +nlat, cases)
        Items are (0-based index): 
        0= TSF=surface temperature at 1 am, 1= TSF at 13 hours,
        2= HEATMM=heat flow, 3= FROST4=frost amount, 
        4= TTB4 = predicted mean bottom temperature
        The prefix contains DJU5 

    55= (seasons,NDX+ items,cases).  For seasonal studies at one latitude
        ITEMS intended to be recoded as needed. Initial version is 9 items:
        [Tsur@ 1am,3am,1pm, spare, Tplan @1am,1pm, Surface heat flow,
        frost budget, T_bottom]
        The prefix contains DJU5        
         Can hold very large number of seasons and cases. 
        THIS MODE DOES NOT SUPPORT CONTINUATION RUNS

    56= [vectors&items, latitudes, NDX+ seasons, cases]
The first dimension is: TSF for all hours, TPF at all hours, 
 T4 for all layers at midnight, then FROST4,HEATMM,TTA4
The prefix is identical to Type 52
\end{verbatim}
Once a disk file is opened, any records written will go into that file until a
new filename is specified (Type 8 Change line), which closes the current file.
It is best to ensure that output file does not already exist. If the file
already exists, new output may be written in same area, even if the old file was
larger than needed.

\subsection{Maximum sizes} %---------------------------
Values for latest version of KRC in this section are in square brackets. All are
firm-coded in krcc8m.f

For any run, even without recorded output, there are three limits:
\qi maximum number of layers, N1 [50]  
\qi maximum number of times of day, N2 [393216]
\qi maximum number of iteration days, N3 [16]

For all file output types, there are two more limits:
\qi The number of latitudes N4 [37] 
\qi  Number of stored hours N24 [96]. 
\\ All of these are checked and limited in TCARD before a run starts.

MAXN5 and MAXN6 are not limits for standard KRC.

\subsubsection{Direct acccess files} %.....
For type 0, -1, -2 and -3 files, each recorded season is a logical record, so
there is no limit to the number of seasons allowed.

\subsubsection{Type 52 files} %.....

For type 52, the primary limit is the total words available to accumulate
results into a bin5 array; this is the parameter KOMMON in krcc8m.f, currently
set at 10,000,000. The size needed for one case is approximately:  
N24*7*N4*(N5-JDISK+3).

All five limits mentioned above are in effect.

The number of cases allowed is set by the size of case one, and printed as MASE
at the end of the first case in the print output. Cases beyond the maximum that
can be stored will be executed, but not saved.

\subsection{Logical units} %_______________________
See table \ref{filecon.tab}. Note that setting IDB1 to 99 will redirect all IOPM
and IOERR output to IOSP so that all print output is chronological there. This
can be usefull in tracking down causes of errors.

\begin{table} 
\caption{Guide to file connections.}
\label{filecon.tab}
\begin{center} \begin{tabular}{ | r l c c l l | } \hline
Type8 & Routine & UNITS & UNITS & FILCOM &   \\
Index & used    & IOD  & flag & name   & Discussion \\  \hline
 3 & TFAR:4     & IOD3=13 & LOPN3 & FFAR & fff Tsurf; must remain open \\
 4 & TFAR ?      & IODA=14  & LFATM  & FFATM  & ttt Tatm;  must remain open \\
 5 & TDISK:6,2,7 & --  & LOPN4 & FDISK  & .type 52 output, calls BINF5 calls primeio, brief \\  
21 & TDISK:1,2,4 & IOD2=2  & LOPN2 & FDIRA & write .tmx, till closed \\
22 & SEASALB  & IOD1=1  & KVALB & FVALB & read, calls READTXT360, brief \\
23 & SEASTAU  & IOD1  & KVTAU=1 & FVTAU & read, calls READTXT360, brief \\
24 & CLIMTAU  & --    & KVTAU=2 & FVTAU & read, calls BINF5 calls primeio, brief \\  
   &          &     & &       & \ \  error could leave FID open \\
25 & READZONE & IOD1=1  &  LZONE & FZONE & read,zone table, brief  \\
- & input & IOIN=3 & - &  & input file. Default is krc.inp or [any].inp \\
- & keyboard & IOKEY=5 & & & manual input \\
- & print & IOSP=7 & - & & print file. default is [any].prt \\
- & errors & IOERR=9 & - &  & eLog[date:time to msec] \\ 
 \hline  \end{tabular} \end{center} \end{table}


\subsubsection{ASCII Output Files} %------------------
Reconstructed  2017oct01
\vspace{-3.mm}
\begin{verbatim}
----.prt  General results. Things in output are controlled by LP1:6 & LPGLOB
                          and most of the six IDBG values

eLog---.---   Error Log, named for time as yyyymmddThhmmss.fff

fort.43, 44 and 47. tfine8.f , Usually only if eclipse case. see eclipse.tex 

fort 71  tlats8.f:  Debug: surface conditions

fort.72  tlats8.f:  Debug: temperatures and frost prediction

fort.75  tlats8.f:  Debug: geometry details

fort.76  tlats8.f: mimic Mike Mellon ASCII files

fort.77    tun8.f:  Special items, firmcoded

\end{verbatim}
\section{Use of logical units} %____

\begin{table} 
\caption[Logical units]{Guide to logical units used.'D' means only if D-lines are compiled }
\label{lunit.tab}
\begin{center} \begin{tabular}{ | l r l l | } \hline
name  & number &  routine[s] &  purpose \\ \hline
IOD1  & 1    &       & explanation file ||  zone table (briefly open) \\
IOD2  & 2    & tdisk & direct-access write ( and read) \\
IOIN  & 3    &       & Input file: defined in units.inc \\
IOKEY & 5    &       & keyboard:     " " \\ 
IOPM  & 6    &       & moniter:      " " Can be redirected \\
IOSP  & 7    &       & print file:   " " \\
IOERR & 9    &       & errors:       " " was IOPM \\
IOD3  & 13   &       & direct-access far-field, read only \\
IODA  & 14   &       & direct-access far-field, atm, read only \\
ifun & \^{ } &  tfar & iod2 or ioda  \\
? &  ? & BINF5 &  primio system, brief \\
- & 43 & tfine & \\
- & 44 & tfine & \\
- & 46 & tday  & \\
- & 47 & tday  & \\
- & 47 & tfine & \\
D & 71 & tlats & \\
D & 72 & tlats & \\
D & 73 & tday  & \\
D & 75 & tlats & \\
- & 76 & tlats & \\
- & 77 & TUN8  & special user firm-code output \\
D & 88 & tlats & \\  \hline
\end{tabular} \end{center} \end{table}

\section{Optional input files} %_______________________
\subsection{Text}
\subsubsection{Zone depth table}
Name is defined by a change card: 8 25 x 'fileName' / . May have up to 20 lines
of comments before a ``C_END'' line. Then each row define a zone and must
contain 4 columns:
\qi Col 1: ZDZ; thickness, m
\qi Col 2: ZDEN; density, kg/m$^3$ \hspace{2.0 cm} But: =-1=use DENS, -2=use DENS2
\qi Col 3: ZCOND; Conductivity, SI Units. \ $\lceil$   If negative then col 4 is a pointer: 1=upper material, 2=lower material
\qi Col 4: ZSPH; Specific heat, SI Units. \ \  $\lfloor$ as in real input 3:16,  +2 for at T-dependent at Tglobal
\\ Must be at least 3 zones; otherwise could use two layers within input parameters. 
\\ The thickness of the last zone is ignored; KRC will fill out to N1 geometric layers.
\\ Must not be any trailing blank lines

The file name is read in by TCARD; TDAY will call READZONE to read the
file. Then the zones will be expanded into an appropriate number of layers. If
the zone definitions violate the convergence stability requirements or the
number of layers assigned, the case will be skipped with a warning.

MAXN1 [firm-coded as 50 in KRCCOM] zones are allowed, any beyond this will be
ignored.

If the number of layers generated exceeds N1, a warning is printed and the case
is skipped.

\subsubsection{Seasonal albedo table}
Name is defined by a change card: `` 8 22 x 'fileName' / `` and stored as FVALB
in FILCOM. May have up to 20 lines of comments before a ``C_END'' line. Then
each row must contain 2 columns.
\qi Col 1: $L_S$ in degrees
\qi Col 2: surface albedo
\\ Values of Col 1 must be within 0:360 and monotonically increasing. 
\qi Must not be any trailing blank lines

Values apply to all latitudes. Active flag is KVALB

The file name is read in by TCARD; TLATS will call SEASALB to initiate or to get
a values for a specific season. SEASALB calls READTXT360 to read the
file. SEASALB will do linear interpolation in season and handle wrap-around.

Example: \nf{-/krc/tes/valb1.tab}

\subsubsection{Seasonal opacity table}
Similar to the albedo table with these exceptions:
\qi primary routine is SEASTAU.
\qi Name is defined by a change card:  8 23 x 'fileName' and stored as FVTAU in FILCOM
\qi Col 2: visual dust opacity
\qi Active flag is KVTAU=1

Example: \nf{-/krc/tes/SmithTau.tab} 

\subsection{Binary: Seasonal climate table}
Name is defined by a change card: 8 25 x 'fileName' / and stored as FVTAU in
FILCOM; name is shared with variable tau as cannot be active at once. The active
flag is KVTAU=2.

A binary file of .bin5 format. [Seasons,latitude,item] assumes Seasons and
latitude on a rectangular, uniformly-spaced, grid. Item 1 is visual opacity.
Item 2 is ice-cloud opacity [could be optional, but coded to be present].

File is read and bi-linear interpolated by CLIMTAU; which is called by TLATS. 
  
Firm-coded for 72 seasons and 36 latitudes, both uniformly spaced. 

Example: \nf{/work1/mars/opacity/THEMIS1yearDustIce.bin5}

\subsection{Far-field files (fff) \label{fff} }
Prior to version 3.4.1, for sloped surfaces the radiation coming from ``below
the horizon'', the ``far-field'', was assumed to be from a surface with the same
temperature and properties as the sloped surface being calculated; this is
called ``self-heating'' And, the atmosphere temperature was calculated as if the
entire surface was at the current slope. Version 3.4.1 introduced the concept of
a far-field file (fff), output from a prior KRC run, that set the far field
surface temperature and was used for the atmosphere temperature.

Version 3.6.1 allows the far-field surface temperature and the atmosphere
temperature to come from different prior runs. Specifically, the atmosphere
would typically be from a ``flat'' run; as commonly the surface regionally, say
over a radius similar at least as great as the atmospheric scale height, is
close to level. The far-field surface viewed by a slope is typically flat or
tilted toward the current slope, i.e., of opposite azimuth.

Reminder: KRC direct-access -n files have records that are a fixed size, based
on the maximum sizes of N24 and N4, and have one record for each temperature
stored , Ts [Tp, [Ta]] (based on the value of K4OUT) for each season.  The
original KRC run input parameters are stored in the first record.

However, only the emissivity is extracted from the first record. The atmosphere
paramters of an fff should match the current run.

FFF's can be written as REAL*8 (the default) or REAL*4, depending on the values
of ID24. When read, the precision is auto-detected and a REAL*4 fff is converted
to REAL*8 in the TFAREAD routine



Type -n direct-access files are written by a change line:
\vspace{-4.mm} 
\begin{verbatim}
8 21 0  'fineName.tmx' / output a direct-access filere
\end{verbatim}
\vspace{-4.mm} 
which lasts only for one case.

Type -n direct-access files to be used as fff are read by a change line
\vspace{-4.mm} 
\begin{verbatim}
8 3 0  'fineName.tmx' / read a fff to be used for far surface
8 4 0  'fineName.tm3' / read a fff to be used for atmosphere
\end{verbatim}
\vspace{-4.mm} 
which is active until turned off by using a similar change line except the entire file name has no more than 4 characters; e.g.,
\vspace{-4.mm} 
\begin{verbatim}
8 3 0  'off' / 
\end{verbatim}

If there is no atmosphere, use is simple. Type -1,-2, or -3 can be used, but only Tsurface is active.

With an atmosphere, it is more complex. An 8 3 ... change line will read a file
to be used for the far surface. If that file is type -3, its atmosphere values
will be used. If not, then an 8 4 ... change line specifying a type -3 file must
be present; such a line overrides any active fff atmosphere values.  Both 8 3 ... and 8 4 ... values remain in effect until turned off by
a corresponding change line with a short file name.
\vspace{-3.mm} 
\begin{verbatim}
8 4 0  'fineName.tmx' / read a fff to be used for the atmosphere
\end{verbatim}

\subsubsection{fff Algorithm Details:} 
All interface with an fff, except closing the unit, is through TFAR.
\\ TCARD:
\qi 8 3 ... closes logical unit IOD3, sets IOPN3 false, and updates the FFAR file name
\qi 8 4 ... closes logical unit IODA, sets LFATM false and updates the FFATM file name
\\ KRC: 
\qi If IOPN3 false and FFAR length 4+, calls TFAR(1 to open FFAR 
\qi If LFATM false and FFATM length 4+, calls TFAR(11 to open FFATM
\qii If that  fff case was sloped, prints warning 
\qi If the only type -3 read is FFAR, uses that for Tatm and sets FAME true (same file for both T's)
\qi If case has an atmosphere, and fff surface is specified but no fff atm is specifed; case will be skipped as invalid.
\\ TSEAS:
\qi Calls TFAR to interpolate to current season
\qii If LOPN3 true, updates Tsurf using FFAR. If fff is type -3, Tatm is also updated. 
\qii If KRC case has an atmosphere and LFAME=false, then updates Tatm using FFATM
\\ TFAR:
\qi Updates IOPN3 or LFATM when it opens the associated file
\qi Returns as many temperatures as are in the file.

\section{Handy things} %_______________________

The first "hour" in printout and output arrays is  1/N24 of a
sol after midnight. E.g., the last time is midnight, not the first.

Atmospheric scale height, SCALEH, depends upon physical constants GRAV [input] 
and TATMAVE which (2007nov) is TATM [input] for the first season and 
thereafter the diurnal average of the prior season. 
 
To run and save various cases for a single season, set N5 and JDISK to 1.

To extract a detailed day by saving DAYCOM to disk, set JDISK=N5, set a new
file name, and set K4OUT to desired latitude index (normally 1):

To run continuously with output every K (1-3) days, set DELJUL=K*PERIOD
This will force prediction terms to near 0.
\qi        setting N3=1 will turn off all prediction.
\qi        set GGT large (to avoid iteration for convergence)
\qi        set NRSET=999 (to avoid reset of layers)

To continue a run with new parameters (e.g., DELJUL)
\qi	3 21 1 'flag set to continue' \\
Note: changing DELJUL will cause reset of DJUL \\
Must increase the value of N5: e.g., 2 5 [bigger] 'Increase stopping season' 
\qi Reset will not occur because J5 continues incrementing


\subsection{ To run two material types } %--------------

Set IC2 to the first layer to have the lower material properties ( $\geq 3$) \\
Set COND2 to the lower material conductivity \\
Set DENS2 to the lower material density \\
Set SPHT2 to the lower material specific heat \\
If LOCAL is False, then initial setting of all layer thicknesses is based
upon the scale of the upper material; if it is set True, the thickness of the
lower layers is set by their scale. \\
  TDAY no longer allows unstable (thin) layers, and will increase the thickness
of the layer IC2 to satisfy the convergence safety factor FCONV if needed. 
However, the code to check on convergence was retained.

\subsection{Setting temperature-dependant properties} %--------------

Basic Flag is L10=LKOFT . If this is true, then the 8 input parmeters ConUp0 to
ConLo3 must be set to yield thermal conductivity as a function of temperature
for the upper and lower materials. $ k=c0 +c_1x + c_2x^2 +c_3x^3 $ where
$x=(T-200.)*0.01$ 

Correspondingly, the 8 input parmeters SphUp0 to SphLo3 must be set for specific
heat
 
One way to generate the coefficients is to run for each of the upper and lower
materials the IDL procedure KOFTOP, which can call all of the
temperature-dependant routines. KOFTOP allows change of its parameters,
including grain radius and pressure, and will print the required parameters
ready for input to KRC.

Below are sample coefficients for thermal conductivity based on Sylvain
Piqueux's numerical model for un-cemented soils; the fit error is $<0.1$\% over
120-320K. Left column is grain radius in micrometers, then the four normalized
coefficients ready for inclusion in a KRC input file, followed by the thermal
inertia at 220K for nominal density and specific heat.
\vspace{-3.mm} 
\begin{verbatim}
 R(mu)         c0        c1        c2        c3      Iner  
    10.     0.008274  0.000735 -0.000376  0.000148    89.8 
    20.     0.012379  0.001280 -0.000629  0.000250   109.9 
    50.     0.021485  0.002647 -0.001201  0.000483   144.7 
   100.     0.032051  0.004528 -0.001874  0.000761   176.8 
   200.     0.046023  0.007569 -0.002743  0.001129   211.8 
   500.     0.068387  0.014075 -0.003874  0.001687   258.2 
  1000.     0.086303  0.021288 -0.004146  0.002099   290.1 
  2000.     0.103743  0.030909 -0.003141  0.002535   318.0 
  5000.     0.127172  0.049907  0.002019  0.003469   352.1 
 10000.     0.149810  0.074734  0.011546  0.004939   382.2 
 20000.     0.185706  0.119913  0.030938  0.007877   425.5 
 50000.     0.283361  0.250283  0.089327  0.016714   525.6 
\end{verbatim}

--------------------------------------------------------------------------------

 \section{RUNNING THE "ONE-POINT" MODE} %_________________ (2002mar08)

A parameter initialization file   \nf{Mone.inp}  is provided. It sets the KRC 
system into a reasonable mode for one-point calculations. Do not change that 
file unless you have read this entire file; caveat plastes (modeler beware).

A line near the end of that file points to a file 'oneA.one' which can contain
any number of one-point conditions. You can replace that name with your own; the
named file is intended to be edited to contain the cases you want; however, it
must maintain the input format of the sample file.

First Line is any title you wish. It must be present. \\
The second line is an alignment guide for the location lines. It must be there.

Each following line must start with an '11 '; this is a code that tells the
full-up KRC that is a one-point line. The next 9 fields are read with a fixed
format, and each item should be aligned with the last character of the Column
title. All items must be present, each line must extend at least to the m in
Azim; comments may extend beyond that, but they will not appear in the output
file. Be sure to have a $<$CR$>$ at the end of the last input line; i.e., no
blank lines!

\vspace{-3.mm} 
\begin{verbatim}
The fields (after the 11) in the one-point input are:
     Ls      L_sub_S season, in degrees
     Lat     Aerographic latitude in degrees
     Hour    Local time, in 1/24'ths of a Martian Day
     Elev    Surface elevation (relative to a mean surface Geoid), in Km
     Alb     Bolometric Albedo, dimensionless
     Inerti  Thermal Inertia, in SI units
     Opac    Atmospheric dust opacity in the Solar wavelength region
     Slop_   Regional slope, in degrees from horizontal
     Azim    Azimuth of the down-slope direction, Degrees East of North.
     Title   From 1 to 20 characters, must not be entirely blank

The two additional columns in the output file are:
     TkSur     Surface kinetic temperature
     TbPla     Planetary bolometric brightness temperature
\end{verbatim}

Try running the binary file first. If that fails, a Makefile is provided to
compile and link the program; simply enter "make krc" and pray. If this fails,
have your local guru look over the Makefile for local dependancies. Suggestions
of making the Makefile more universal are welcome.

To run the program, change to the directory where the program was built, and
enter "krc". You should get a prompt:
 \qi      ?* Input file name or / for default = Mone.inp  \\
If the initialization file still has this name and is in the same directory,
enter a single "/" and $<$CR$>$. Otherwise, enter the full pathname to the 
initialization file, with no quotes and no blanks.

A second prompt is for the name of the output file: 
 \qi         ?* Print file name or / for default = krc.prt \\
Again, if this is satisfactory, simply enter  / $<$CR$>$ , else enter the desired
file path-name.

\subsubsection{ Comments on the One-point model} %............ 

The initialization file \nf{Mone.inp} is a compromise between the secular trend
being reached (less layers) and the annual variation being damped (more
layers). The version after KRC v2.2.4 was tested over a large grid of
non-polar-cap points against a model twice as deep and run for 6 years; the
average absolute difference was 0.05 K. Execution time on a circa 2013 PC is
about 30 cases/second.

The underlying model is the full version of KRC. By modifying the initialization
file, you can compute almost anything you might want. If you choose to try this,
best to read all of this document.

\section{ DEBUG OPTIONS \label{debug}}
If the first input line has a no-zero third number, then the second line is 6 
white-separated debug-control integers: IDB1 to IDB6
% grep -in idb *3.f     grep -in idb tseas8.f

Most of these were convered into D-lines in v361, except the IDB2.GE.3 drint of zone table details in TDAY

\vspace{-3.mm} 
\begin{verbatim}
2018 Oct 23 01:59:14  grep -in IDB *8.f > qdb  then   grep -i idb1 qdb  etc

porb08.f:16:        LOP=IDB1.GT.3
tcard8.f:128:     IF (IDB1.EQ.99) THEN 'Switching IOPM and IOERR output to IOSP'
tcard8.f:155:      IF (IDB1.GE.1) WRITE(IOPM,*)'Before PORB0'
tcard8.f:157:      IF (IDB1.GE.1) WRITE(IOPM,*)'AFTER PORB0'
tcard8.f:459:      IF (IDB1.NE.0) WRITE(IOSP,*)'TCARD Exit: IRET=',IRET,NFD,ID(1)
tseas8.f:59:      IF (IDB1.NE.0) WRITE(IOSP,*) 'TSEASa',IQ,J5,LSC,N5,LONE

tcard8.f:102:      IF (IDB2.GE.5) WRITE(IOSP,*) 'TCARD-A',IQ
tday8.f:87:      IF (IDB2.GE.5) WRITE(IOSP,*) 'TDAY IQ,J4=',IQ,J4,JJO
tday8.f:160:        IF (IDB2.GE.3) WRITE(IOSP,*)'Zlay  K  J   DSCAL     DDZ'
tday8.f:233:          IF (IDB2.GE.3) WRITE(IOSP,'(a,2i3,6f8.3,i3,f8.4)')  
tday8.f:599:      IF (IDB2.EQ.2) WRITE(IOSP,119) LZONE,LALCON,j5,IK1,IK2,IK3,IK4
tday8.f:885:      IF (IDB2.EQ.2) WRITE(IOSP,119) LZONE,LALCON,j5,IK1,IK2,IK3,IK4
tday8.f:905: 9    IF (IDB2.GE.6) WRITE(IOSP,*) 'TDAYx'
tlats8.f:52:C 2016sep09 HK Do not print T-equilb unless IDB2.ge.5
tlats8.f:129:      LQ1=IDB2.GE.5             ! once per season or latitude
tlats8.f:130:      LQ2=IDB2.GE.9             ! each time of day
tlats8.f:131:      IF (IDB2.NE.0) WRITE(IOSP,*)'TLATSa',N3,N4,J5,LATM,LQ1,LQ2
tlats8.f:274:          IF (IDB2.GE. 5) then
tlats8.f:291:            IF (IDB2.GE. 5) then
tlats8.f:647:          IF (IDB2.GE.1) WRITE(IOSP,*)' Tatm,Beta=',TATMJ,BETA
tlats8.f:667:      IF (IDB2.EQ.4) WRITE(IOPM,*)'J4,5 +',J4,J5,TEQUIL,TATMJ
tlats8.f:775: 9    IF (IDB2.GE.3) WRITE(IOSP,*)'TLATSx',N1,N1PIB,N2,N24,J3

tdisk8.f:123:      IF (IDB3.GE.3) WRITE(IOSP,31)'TDISKa',KODE,KREC,J5,K4OUT,ITOP
tdisk8.f:124:      IF (IDB3.GE.4) WRITE(IOSP,31)'TDISKb',N3,N4,N5,J5,MASE !<dbug
tdisk8.f:153:      IF (IDB3 .GE. 4) THEN
tdisk8.f:162:      IF (IDB3 .GE. 4) THEN
tdisk8.f:318:        IF (IDB3.GE.3) WRITE(IOSP,*)'TDISKc KREC=',KREC,IOD2,ITOP,I
tdisk8.f:376:        IRET=IDB3  !  +1=report many values   +2 report progress 
tdisk8.f:393:      IF (IDB3.GE.7) WRITE(IOSP,*)'TDISKx  KREC,JREC=',KREC,JREC
tfine8.f:541:        I=IDB3     !  +1=report many values   +2 report progress 
tlats8.f:762:        I=IDB3     !  +1=report many values   +2 report progress 

krc8.f:122:C- Lines are to redirect error to monitor, but IDB4 not yet known!
krc8.f:125:C-      IF (MODULO(IDB4,2) .EQ. 1) THEN ! no error file 
tday8.f:389:      IF (IDB4.GE.2) WRITE(*,'(a,2I4,2f8.5,f10.3,e12.5)') 
tday8.f:407:        IF (IDB4.GE.2) WRITE(IOSP,'(i3,2f12.3)') J,SCONVG(J),ZDZ(J)
tday8.f:715:D            IF (IDB4.EQ.4 .AND. MOD(JJ,NZ).EQ.0 )  ! N48 per day
tday8.f:828:D           IF (IDB4.EQ.4) WRITE(73,741)J5,J4,JJJ,-1,SNOW,EFROST,TATMJ
tlats8.f:566:C         IF (J5.EQ.IDB4) WRITE(53,531)J4,JJ,TATM,AVET,HUV ; 2018jun

tday8.f:666:              IF (IDB5.GE.4)WRITE(47,22)(TTJ(I),I=1,N1) ! coarse  T
tday8.f:798:          IF (JSW.GT.0 .AND. IDB5.GE.7) WRITE(46,244) 
tfine8.f:107:      IF (IDB5.GE.1) WRITE(IOSP,*) 'TFINE IQ,J4,J5=',IQ,J4,J5
tfine8.f:139:      IF (IDB5.GE.1) WRITE(IOPM,*)'QB..',QB,RLAY,RLAF,BF1,N1F,JLOW
tfine8.f:155:        IF (IDB5.GE.6) WRITE(IOPM,30),'I,j..',I,J,QB,DIFFI(I)
tfine8.f:201:      IF (IDB5.GE.1) WRITE(IOSP,'(a,i3,f12.3)') 
tfine8.f:225:        IF (IRET.LT.1 .OR. IDB5.GE.2) WRITE(42,'(A,I4,2G12.5,F8.1)')
tfine8.f:340:      IF (IDB5.GE.4) THEN
tfine8.f:357:      IF (IDB5.GE.3) WRITE(IOSP,'(a,2f12.4,g13.5)') 
tfine8.f:372:      IF (IDB5.GE.2) WRITE(IOSP,119) LZONE,LALCON,j5,IK1,IK2,IK3,IK4
tfine8.f:478:          IF (IDB5.GE.6 .AND. (JJ.LT.(J7P+3)  .OR. ABS(JJ-J8).LT.3))
tfine8.f:515:      IF (IDB5.GE.4) THEN
tfine8.f:551: 9    IF (IDB5.GE.1) WRITE(IOSP,*) 'TFINE exit'

tfar8.f:87:      IF (IDB6.GE.2) WRITE(IOSP,*)'TFAR:0 ',KODIN,FFATE,FFOLE
tfar8.f:104:      IF (IDB6.GE.2) WRITE(IOSP,*)'TFARfile: ',FILEIN(1:ILEN),I
tfar8.f:292:      IF (IDB6.GE.4) WRITE(IOSP,*)'TFARx:FFOLE+ ',FFOLE,IFUN,LOPF
tlats8.f:187:      IF (IDB6.GT.3) WRITE(IOPM,*)'SKYFAC,FAC5X=',
tseas8:129        IF (IDB6.GE.5) WRITE(IOSP,130) I1,I2,FARTS(I1,I2,1)
about 2017    grep -in lq tlats33.f
128:      LQ1=IDB2.GE.5             ! once per season or latitude
129:      LQ2=IDB2.GE.9             ! each time of day
130:      IF (IDB2.NE.0) WRITE(IOSP,*)'TLATSa',N3,N4,J5,LATM,LQ1,LQ2
131:D       WRITE(IOPM,*)'TLATSa',N3,N4,J5,LATM,LQ1,LQ2 !<dbug
175:        IF (LQ1) I=3
197:      IF (LQ1) THEN
         WRITE(75,*) 'J5+',J5,SUBS,SDEC,DAU,SLOPE,SLOAZI
         WRITE(75,*) 'MXX+',MXX,SKYFAC,FAC5X
241:      IF (LQ1) WRITE(IOPM,*)'TLAT1 J5,TBLOW=',J5,TBLOW
    243:top of Lat loop
247:      LQ3=LD19 .AND. (J5.EQ.JDISK) .AND. (J4.EQ.1) ! first recorded season
270:          IF (LQ1) KODE=4 ! do debug print in  CUBUTERP8
325:      IF (LQ1 .AND. J5.EQ.1) THEN
         WRITE(75,*)'FXX+',FXX,J4,DLAT
         WRITE(75,*)'QXX=',QXX 
         WRITE(75,*)'TXX=',TXX
         WRITE(75,*)'PXX+',PXX,LPH,PARW(7),COSP
409:      IF (LQ1) WRITE(IOPM,*)'TLATS: J4+.',J4,SOLR,ACOSLIM,COSIAM(1)
411:C3      IF (LQ3) WRITE(IOSP,701)'LQ3',NCASE,J5,J4,TATMAVE,PRES,OPACITY
   431  top of time loop
529:         IF (LQ2.AND.(MOD(JJ,24).EQ.1)) THEN
           WRITE(75,*)'HXX+',HXX,JJ
           WRITE(75,*)'ANG:',ANGLE,COSI,COS2,DIRECT,QI
533:         IF (LQ2) WRITE(IOSP,*),'TLatc',JJ,COSI,COS3,DIRECT,DIFFUSE 
548:C3         IF (LQ3) WRITE(88,777)JJ,COSI,COLL,HUV,QI,DIRECT,DIFFUSE,BOUNCE
551:          IF (LQ3) WRITE(88,777)JJ,COSI,COS2,AVET,HALB,DIRECT,ATMHEAT
    559: end time loop
582:          IF (LQ1) WRITE (IOPM,*) 'QS, small tau=',QA,QS
602:      IF (LQ1) then
        WRITE(IOPM,*)'AVEA...',AVEA,AVEE,AVEI,AVEH
        WRITE(IOPM,*)'CABR...',CABR,TAUD,TAUIR,FACTOR,TAUEFF
        WRITE(IOPM,*)'BETA...',BETA,QS,SIGSB 
        WRITE(IOPM,*)'TAEQ4,TSEQ4,TEQUIL',TAEQ4,TSEQ4,TEQUIL
618:        IF (LQ1) WRITE(IOPM,*)'TSUR,TBOT',TEQUIL,TSUR,TBOT 
620:        IF (LQ1) WRITE(IOPM,*)'XCEN',XCEN 
634:          IF (LQ1) WRITE(IOPM,*)'Radiation time, sec',QA 
652:      IF (LQ1) WRITE(IOPM,*)'TTJ',TTJ
    759:  end of lat loop
761:      IF (LQ1 .AND.(.NOT. LONE) .AND. (J5.LE.1)) ! avoid line in  OnePoint output
     &  WRITE(IOSP,*)'TLATS: TEQQ',(TEQQ(I),I=1,N4) ! starting  Tequil 
end lat loop
\end{verbatim}

\section{Accessing KRC output in IDL} %______________________

IDL routines do not access files directly unless specifically listed.

DEFINEKRC \ Define structures in IDL that correspond for Fortran commons \\
Calls: None == None other than IDL library \\
Firm code of common definitions. Must be recoded if a Fortran *.inc changes

KRCSIZES \ Compute array and common sizes for KRC Fortran \\
Test procedure to compute array sizes or hours. \\
Must recode if any size in *.inc changes \\
Calls: None

READKRCCOM \ Read a KRCCOM structure from a bin5 file \\
Uses 3-element HOLD array. Returns a structure of krccom \\
Options to open or close bin5 file or read one case  \\
Calls: DEFINEKRC \\
Files: bin5 \\
HOLD is: 0]=logical unit  1]=number of words in a case  2]=\# cases in the file 

READKRC1 \ Read KRC direct access filesC \\
Calls:  DEFINEKRCC \\
Read type -1,-2 or -3 files. Caller sets the KRC version that wrote the file. 

READKRC52 \ Reads a KRC type 52 bin5 file \\
Calls:  BIN5read  READKRCCOM \\
Reads and unpacks type 52 file, returning 5 multi-dimensional arrays

KRCHANGE   Find changes in KRC input values in common KRCCOM \\
Calls:  READKRCCOM  MAKEKRCVAL \\
Reads and stores krccom for first case. For each additional case, makes a 
list of any changes in the float, integer or logical input values. 

KRCCOMLAB \ Print KRC common input items \\
 all items via arguments \\
Calls: None

MAKEKRCVAL \ Make string of selected KRC inputs: Key=val \\
Calls: DEFINEKRC

KRCLAYER \ Compute center depth of KRC layers \\
 all items via arguments \\
Calls: None

\input{tun}  %<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<

\section{Notes on how some aspects of the code work} %___________
Dates after some subsection titles indicate last update.

\subsection{New file name}%----------- 
TCARD reads a card of Type 8, (and index is not 22 or 23)
\qi it calls  TDISK(4,0), which closes current file and sets  LOPN2=.FALSE.
\qii   TCARD then moves new file name into common \\
KRC checks if current (new) values of N5 and JDISK call for file output;
\qi  with  LOPN2=.FALSE., KRC calls CALL TDISK (1,0) to open new file.

\subsection{End of a case and end of a run}%-----------
TCARD sets KOUNT=0 at entry; this is incremented for every card except those of
type 0 ( or less) or type 11 (one-point mode). When type 0 is encountered, if
KOUNT is positive, does normal check of changes before return with IR=1 to
indicate start of a new case; if KOUNT is zero, returns with IR=5 and prints
'END OF DATA ON INPUT UNIT'

\subsection{Setting one-point mode}%-----------
This can be done only in the first case, and there is no way to leave the 
one-point mode except to end the run.

TCARD encounters: " 10 * filename" as change card in the initial case.
\qi   sets this as new input file name, then returns with IRET=4 
\qi  [Thus, nothing following this change card in initial file is read] \\
KRC closes prior input file, opens the new one, and reads past first two lines 
\qi     then calls TCARD to read first one-point line and sets LONE=true
\qi     and drops into the top of the "case" loop. \\ 
The master one-point should have a single latitude, no binary output file. \\
The small number of layers, days to converge, and seasons ignores the seasonal 
effect.

One-point request values are read by TCARD @ 310, which computes starting DJUL

TPRINT does linear interpolation of TOUT, which has N2 points be sol. To get Tp,
does interpolation of Tp-Ts at the hour points, and adds to interpolated Ts.
 \subsubsection{How one-point converts Ls to date}
 Ver 212: 
XREAD is the 2nd column in the OnePoint file, i.e., Ls. 

In TCARD 310: calls PORBIT to get the date of the desired Ls, then backs up
(N5-1)*DELJUL to the starting date.

\subsection{Starting conditions and date}%-----------

Initial N5-JDISK sets the size of output files. There could be any number of
interior seasons where parameter changes are made; based on successive values of
IDOWN.

KRC initially calls TCARD(1  \\
For each case loop, sets IQ=TCARD\_return. If one-point mode, sets IQ=1

TSEAS uses IQ as key. It this is 1, then sets J5=0 and sets DJU5 to season -1.,
else, increments J5 and increments DJU5 with current DELDUL. This allows use of
variable resolution dates. (so J5 never 0 when TCARD(2 called) \\

TLATS uses J5 as the key; if it is $<= 1$, then starts from equilibrium
conditions, else uses predictions from prior season

The default is that change cards cause a fresh calculation of starting
conditions. Exceptions are when J5=IDOWN$>0$ at TCARD entry

\subsection{Changing parameters within a seasonal run = Continue from memory.}%-----------

When J5 reaches IDOWN, TSEAS calls TCARD, which will set IRET=3 before reading
the new parameters. May change DELJUL to get finer seasonal resolution, but must
NOT change N5

Use: Normal restrictions for what may not change for Type 52 files apply.
% E.g., type 56 must NOT change number of latitudes nor total number of seasons.

Set N5 to be the total number of seasons desired, including those
after any number of parameter changes; it must NOT be changed later.

Set IDOWN to the season at the beginning of which wish to (first) change
parameters. The next set of changes could include a revised (larger) IDOWN.

CAUTION: This has not been tested in recent versions 


\subsection{Use of common PORBCM}%-----------
Contents are described in porbcm8.f \\
PORBCM is filled by TCARD calling PORB0, which reads the first 30 items in 
5G15.7 from the input file and sets the value of $\pi$ and radians-to-degrees. 

TSEAS call PORBIT to get Ls, the heliocentric range and the sub-solar latitude.

\subsection{Lower boundary condition and resetting (jumping) layer temperatures.} %-----------
At the start of a case, TLATS sets the temperature profile linear with depth
in one of three ways:
\qi IIB=0: top and bottom at equilbrium temperature
\qi IIB=1: top at equilbrium temperature, the bottom at TDEEP
\qi IIB=2: top and bottom at TDEEP

The kind of resetting is controlled by IB. In TCARD, if IB$>$0, then N1PIB=N1+1,
else N1PIB=N1.  T(N1+1) is not reset in the time calculations. In TDAY, for each
time step, the temperature of the lower boundary is set equal to T(N1PIB), which
results in either zero heat flow (IB=0) or a constant temperature.

\subsection{Temperature scaling for KofT} 
\nv{TOFF} and \nv{TMUL} are firm-coded in \nf{tday} as 220 and 0.01. These are
appropriate for Mars.  They may be made input parameters in later versions to
address Mercury-like planets. As of v3.5.1 the scaling is firm-coded in
EVMONO3D.

\subsection{Atmosphere flag} PTOTAL is an input parameter and in KRCCOM. 
\np{tseas8.f, tlats8.f, tday8.f} and \np{tfine8.f} each contain:
LATM=PTOTAL.GT.1. \ so a pressure of 1.0Pa or less is treated as zero
atmosphere. The \nv{LATM} flag is used liberally in those routines.  \np{tday}

\subsection{ Cap-dependent pressure 2018jun22} %-----------
If there is an atmosphere and more than 9 latitudes (which should be ordered
from south to north), KRC will compute the global average surface frost (SUMF,
kg/m$^2$) each season by TSEAS calling TINT8; otherwise, SUMF=0. If the input
KPREF is 2, then in TLATS SUMF is converted to the equivalent partial pressure
and subtracted from the input total pressure.

\subsection{Frost condensation temperature 2018jun22 } 
The current frost condensation temperature is TFNOW (in KRCCOM). If there is an
atmosphere and the LVFT input flag is false, then TFNOW is set to the TFROST
input parameter; if LVFT is true, then TFNOW is reset each season and latitude
using the routine GASPT based on the local current surface partial-pressure of
the condensing gas (PFACTOR*PGASG) and the two Clausius-Clapeyron input
parameters. The atmospheric (cloud) condensation temperature is based on
PFACTOR*PGASG /$e$

\subsection{Far-field temperatures 2016jun30}
FFAR initially set by KRC to be 'no'; this is less than 4 characters, so
far-field is considered turned off.

An input card ' 8 3 x text': TCARD loads $<$text$>$ into FFAR. 
For each case, KRC then checks length of FFAR; if 4 or more, will call TFAR(1 
to open a FAR file for read. If successful, TFAR sets LOPN3=.true. 
\\ LOPN3 is used in TDAY as the flag that the far-field calculation is active
(as opposed to self heating)

\subsection{ Seasonal variation of albedo or opacity or ``climate''}%-----------

When TCARD encounters a type of 8 and an index of 23(tau) [or 22(albedo)], it
transfers the text item into FVTAU (which is in COMMON /FILCOM/) and then calls
SEASTAU with an Ls of -999 .  SEASTAU when called with LSUB LT -90 calls
(providing IOD3) READTXT360, which reads file. Maximum number of rows is 360,
more will be ignored. First and last entry read are wrapped with $\pm$360 to Ls
to ensure no interpolation faults later. TCARD sets the variable Tau flag, KVTAU,
to 1 if table-read was successful, else it is set to 0.
If KVTAU is set to 1, TSEAS calls SEASTAU at start of each season, resetting TAUD. 

If type 8 and index 23, the same as above except names are -ALB rather than -TAU

If type 8 and index 24, then TCARD calls CLIMTAU to open and read a .bin5 
climate file, and sets KVTAU=2

CLIMTAU expects to read a .bin5 file (season,latitude,2) with dust and ice
infrared opacities; this file is normally made by the IDL routine mopacity.pro .
Season is assumed to the uniform in Ls from 0 to 360-delta and latitude assumes
to be uniform from -90+delta/2 to +90-delta/2. CLIMTAU has firm-coded sizes, 72
seasons and 36 latitudes, and stores the file. Upon later calls from TLATS, it
returns the two opacities at a requested Ls and latitude, using bi-linear
interpolation.

\subsection{ When parameter and layer tables appear} %-----------
The flag LD18 means: a value has changed after the parameter and layer tables
were last are printed. It is set: initially and by TCARD. It is cleared when the
tabel are printed,

 Can tell if zones are active by length of FZONE in common

These two tables are printed: initially, before TSEAS for each case if LP2 is
set, or after TSEAS if an error occurs or if LD15

When to print.
\qi first case
\qi if LP2 true
\qi if an error and prints are out of date.
\qii Something is out-of-date: initially, or if TCARD was called
\qii but not for each season, LSC true.
\\ What to print
\qi parameters
\qi layers
\qi zones if active.

If all the prints were at the same place, or controlled from the same place,
that could set the ``current'' flag.

As of 3.3.1, 
\qi parameters printed by TPRINT(2
\qi layers printed in TDAY(3
\qi zones printed by READZONE in TDAY(1 or 3, if active

Because case change is done in KRC8, and errors return there with a code,
 the top control logic could be there.  

\subsection{Atmosphere condensation and snowfall} %-----------.

 Each midnight, the atmospheric temperature $T_a$ is compared to the saturation
 temperature $T_{sat}$ TATMIN computed at the beginning of each season and
 latitude based on the two input Clausius-Calperyon parameters and the partial
 pressure of condensible gas at one scale height above the local surface. The
 local surface pressure $P_s$ (Pascal) is derived from the current 0-elevation
 surface pressure PZREF, the fraction of condensible gas, the local elevation
 and the current local scale-height SCALEH. The transfer of snow from atmosphere
 to ground (surface frost) is considered instantaneous. Prior to version 356,
 the negative energy of snow which occurs when there is no surface frost was lost
 from the system, but recorded as FLOST; this was rare.

 The energy required to warm the atmosphere is $E= (T_{sat}-T_a) \cdot c_a
 P_s/g$ where $c_a$ is specific heat at constant pressure of the atmosphere
 (J/kg/K)and $g$ the surface gravity; the terms after the dot are combined into
 CPOG. The snowfall amount is $E/L_f$ (Kg/m$^2$) where $L_f$ is the latent heat of
 sublimation of frost, input parameter CFROST.

 In a test with latitudes every 5\qd and thermal inertias of 100, 60 and 20,
 loss occurred on 3.5\% of the snowy days with a average loss of 0.63 kg/m$^2$,
 equivalent to 3.7E5 J/m$^2$. The top physical layer in these models was 3.0,
 1.8, and 0.6 kg/m$^2$, respectively, so they would be cooled by 190, 320, and
 960 K, far greater than needed to reach frost temperature.

Thus, beginning in version 3.5.6, the snow is assumed to become surface frost
and the surface set to frost temperature.

%        TATMIN = GASPT(1,PFACTOR*PGASG/2.71828) ! frost point for 1-layer atm
% SNOW= (TATMIN-TATMJ)*CPOG/CFROST ! | snow formation  Kg/m^2
% CPOG=ATMCP*(DMAX1(PRES, 1.D0)/GRAV) ! atmosphere:  D_Energy / d_T
% ATMCP is input and in KRCCOM, PRES is computed and in KRCCOM 
% 48  AtmCp   Specific heat at constant pressure of the atmosphere [J/kg/K]
% 72  PRES    Local surface pressure at current season. compute in TLATS

\subsection{Seasonal prediction \ \  Update 2020apr11} %-----------

Several quantities are stored in TDAY at midnight;
 with index J3P1: TT1,TTS, TTB, and if Atm., TTA,FRO. These are effectively at the start of day J3+1. Index 1 is set initially as the global equilibrium T.; after season 1, these are set set to the prediction from the prior season.
After convergence, TLATS uses EPRED to predict these to the end of the season.
\\ 355 and earlier ALERT:  TTA,FRO not set for index 1.

Up to version 355:
\\ JE is the index  of the first (of three) midnight given to EPRED
\\ FP is the number of sols forward to predict.
\qi Normally JE is J3P1-2 == J3-1
\qi If the prediction is less than 0.1 sol, use force EPRED into linear mode.

Version 356: The full daily time series for a season is passed to EPRED, along with the index of the last valid value.  If 3 or more points valid, does asymptotic predict limited by the latest slope, there is no minimum FP.
Else, if 2 points valid, does linear predication; else, uses the last valid value.

Outline of code: 
\qb For each latitude, if the first season, TTJ is set linear with depth between TSUR and TBOT; which in turn depend upon how IIB is set; otherwise start with forecast for the last midnight of the prior season.
\qb TDAY runs until convergence, saving the profile each midnight.
\qb For each layer, the temperatures at midnight are forecast to the end of the season.

Annotated code extracts:
\vspace{-3.mm} 
\begin{verbatim}
 daycom:: TTJ(MAXN1P)       ! Layer temperatures (TTJ(1) is surface temperature)
          TT1(MAXN1,MAXN3)  ! Temperatures at start of day for each layer and day
 latcom:: TMN4(MAXN1,MAXN4) ! predicted convergence midnight temperature

Line in TLATS [2020 Apr 11 17:39:42]
    before TDAY
642. (J5.LE.1): CALL MVD(TTJ,TT1,N1) ! temperature profile, start of first day
660.   else:    CALL MVD(TMN4(1,J4),TT1,N1PIB) ! copy forecast midnight values 
667. CALL TDAY
715. I=1,N1  
    CALL MV21D (TT1(I,1),MAXN1,WORK,J3P1) ! transfer all defined values for layer I
    TMN4(I,J4)= EPRED8(WORK,FP,J3P1, TFNOW,TBLOW) ! forecast to end of season
\end{verbatim}  


\section{Non-standard FORTRAN}
All thought to be removed in Version 3+

Although the SAVE option is used in some routines, this has not been done for all routines that depend upon values being consistent between successive calls, so `` -fno-automatic `` must be invoked in the Makefile.

\subsection{Modifying size parameters}
Most arrays are sized using PARAMETER's set in krcc8m.f. Hopefully, these could
be modified without breaking the KRC system. If you do so, extensive testing is
recommended.

In particular, if a user were generating jillions of direct access files, it
would be efficient to set MAXN4 and MAXNH to the values of N4 and N24; however,
their product must be larger than the size of KRCCOM (NWKRC), or twice as large
if writing single-precision files. In version 3.6.1 there is a test for this,
which will STOP if the condition is violated.

\appendix %=================================================================

\section{Map of type 52 file \label{type52}}
\input{type52map}

\section{Time-step doubling}
In the JGR article, a convergence safety factor is defined as $ B_i / \sqrt{ 2
  \Delta t_i \cdot \kappa_i}$ ; KRC[103]. In the code, the square of this, 
 $B_i^2 / (2 \Delta t_i \cdot \kappa_i)$ is defined as SCONVG and printed.

Proceeding from the top, if SCONVG is more than twice CONVF for a layer, then
the time-step can be doubled for that layer.

Code changed significantly when layer tables added; between 
\qi v321: layer thickness BLAY set by FLAY (virtual) and RLAY 
\qi and 331: BLAY set by FLAY (top physical) and RLAY
\\ With zone table, at each layer must know the minimum pre-doubling safty factor at all lower layer

For same layer thicknesses, after first time doubling, 321 and 341
converg. factor disagree!

\subsection {Version 343}

Code includes several loops for setting layer parameters, starting at the
indicated line number ( 2016 sept 5 source code):

Set the thermophysical properties for each layer. Seperate exclusive loops for
\qi  156:  zone table or not.
\qi  278: no zone table

327: Print temperature-dependent properties for up to two zones 

369: If not a zone table, check lower material stabilty and increase layer
thickness if needed.

379: LOOP 1 compute safety factor for each layer before any doubling SCONVG(J)

387: LOOP 2. Find the minimum safety factor at or below each layer before any
time doubling; ZDZ (temporary)

406 LOOP 3 Assign time doubling, from top to depth. In English: 
\qi IF:  have not yet reach maximum number of time doublings allowed 
\qi AND at least 2 physical layers already 
\qi AND number of time steps now is even
\qi AND new time-step would be acceptably stable for all lower layers
\\ THEN, do doubling at current layer and update forecast factor.

436: Generate the last-layer versus time comb KJ

480: Print the layer table

\subsection {TDISK called by}
\begin{verbatim}

krc8.f:228:   IF (.NOT.LOPN2 .AND. I.GE.4) CALL TDISK8 (1,0) ! open output disk file   
tcard8.f:137:   CALL TDISK8 (1,1)         ! open old file to read starting conditions

tday8.f:902:  IF (J5.GE.JDISK) CALL TDISK8(2,I) ! write current season if valid
tseas8.f:169:  IF (J5.GE.JDISK) CALL TDISK8 (2,KREC) ! write this season

tcard8.f:138:   CALL TDISK8 (3,KOLD)      ! read starting record; will override  KRCCOM

krc8.f:231:   IF (.NOT.LOPN4 .AND. I.GE.4) CALL TDISK8 (6,0) ! open output type 5x
krc8.f:313:   IF (LOPN2) CALL TDISK8 (4,KREC) ! close FDIRA, cannot hold multiple cases
krc8.f:334:   IF (LOPN2) CALL TDISK8 (4,KREC) ! all done: close disk files
tcard8.f:263: IF (LOPN2) CALL TDISK8 (4,0)    ! close prior direct-access file
tcard8.f:310: IF (LOPN2) CALL TDISK8 (4,0)   ! must close disk files if open
tcard8.f:140:            CALL TDISK8 (4,0)       ! close 'starting' file

tseas8.f:168:  .AND. K4OUT.LT.0) CALL TDISK8 (5,KREC) ! must follow tday(1)
 
krc8.f:231:   IF (.NOT.LOPN4 .AND. I.GE.4) CALL TDISK8 (6,0) ! open output type 5x

krc8.f:337:   IF (LOPN4) CALL TDISK8 (7,KREC) ! type 5x
tcard8.f:258: IF (LOPN4) CALL TDISK8 (7,0)    ! close current bin5
tcard8.f:312: IF (LOPN4) CALL TDISK8 (7,0)    

\end{verbatim} 
\end{document} %===============================================================
%============================================================================
%============================================================================


5 latitude 40 seas with  mean Mars zonal elevations. Effect of atmos.  
KRCv3.2.1    RUN-CASE  1- 1   2016 Jul  1 06:06:38  PAGE=  3
 Conductiv.= 3.864E-02  Dens*Cp= 1.035E+06  Diffu.= 3.733E-08  Scale= 3.248E-02
         ___THICKNESS____    __CENTER_DEPTH__    Total  Converg.
 LAYER    scale     meter     scale     meter   kg/m^2  factor         factor
                                                         SCONVG(I) SCONVG(I)
    1    0.1800    0.0058   -0.0900   -0.0029     0.000   0.000   0.000
    2    0.2160    0.0070    0.1080    0.0035    11.224   2.851   2.851
    3    0.2592    0.0084    0.3456    0.0112    24.693   4.106   4.106
    4    0.3110    0.0101    0.6307    0.0205    40.856   5.913   2.956 2
    5    0.3732    0.0121    0.9729    0.0316    60.251   4.257   2.129 2
    6    0.4479    0.0145    1.3834    0.0449    83.525   6.130   3.065 2
    7    0.5375    0.0175    1.8761    0.0609   111.455   4.414   2.207 2
    8    0.6450    0.0209    2.4673    0.0801   144.970   6.356   3.178 2
    9    0.7740    0.0251    3.1768    0.1032   185.188   4.576   2.288 2
   10    0.9288    0.0302    4.0282    0.1308   233.449   6.590   3.295 2
   11    1.1145    0.0362    5.0498    0.1640   291.363   4.745   2.372 2
   12    1.3374    0.0434    6.2758    0.2038   360.860   6.832   3.416 2
   13    1.6049    0.0521    7.7469    0.2516   444.256   4.919   2.460 2
   14    1.9259    0.0625    9.5123    0.3089   544.332   7.084   3.542 2
   15    2.3111    0.0751   11.6308    0.3777   664.422   5.100   2.550 2
   16    2.7733    0.0901   14.1730    0.4603   808.531   7.344   3.672 2
   17    3.3279    0.1081   17.2235    0.5594   981.461   5.288   5.288
   18    3.9935    0.1297   20.8842    0.6783  1188.977   7.615   7.615
   19    4.7922    0.1556   25.2771    0.8209  1437.997  10.965  10.965
   20    5.7506    0.1868   30.5485    0.9921  1736.820  15.790  15.790
0Bottom layers for time doubling:      4    6    8   10   12   14   16   20

 2016 Jul  2 06:32:21 
         ___THICKNESS____    __CENTER_DEPTH__  Conductiv. Density Sp.Heat   Total Converg.
 LAYER  D_scale     meter   D_scale     meter      W/m-K   kg/m^3   J/kg   kg/m^2  factor
    1    0.1800    0.0058   -0.0900   -0.0029 0.3864E-01  1600.00  647.00     0.000   0.000
    2    0.2160    0.0070    0.1080    0.0035 0.3864E-01  1600.00  647.00    11.224   2.851
    3    0.2592    0.0084    0.3456    0.0112 0.3864E-01  1600.00  647.00    24.693   4.106
    4    0.3110    0.0101    0.6307    0.0205 0.3864E-01  1600.00  647.00    40.856   2.956
    5    0.3732    0.0121    0.9729    0.0316 0.3864E-01  1600.00  647.00    60.251   2.129
    6    0.4479    0.0145    1.3834    0.0449 0.3864E-01  1600.00  647.00    83.525   3.065
    7    0.5375    0.0175    1.8761    0.0609 0.3864E-01  1600.00  647.00   111.455   2.207
    8    0.6450    0.0209    2.4673    0.0801 0.3864E-01  1600.00  647.00   144.970   3.178
    9    0.7740    0.0251    3.1768    0.1032 0.3864E-01  1600.00  647.00   185.188   2.288
   10    0.9288    0.0302    4.0282    0.1308 0.3864E-01  1600.00  647.00   233.449   3.295
   11    1.1145    0.0362    5.0498    0.1640 0.3864E-01  1600.00  647.00   291.363   2.372
   12    1.3374    0.0434    6.2758    0.2038 0.3864E-01  1600.00  647.00   360.860   3.416
   13    1.6049    0.0521    7.7469    0.2516 0.3864E-01  1600.00  647.00   444.256   2.460
   14    1.9259    0.0625    9.5123    0.3089 0.3864E-01  1600.00  647.00   544.332   3.542
   15    2.3111    0.0751   11.6308    0.3777 0.3864E-01  1600.00  647.00   664.422   2.550
   16    2.7733    0.0901   14.1730    0.4603 0.3864E-01  1600.00  647.00   808.531   3.672
   17    3.3279    0.1081   17.2235    0.5594 0.3864E-01  1600.00  647.00   981.461   5.288
   18    3.9935    0.1297   20.8842    0.6783 0.3864E-01  1600.00  647.00  1188.977   7.615
   19    4.7922    0.1556   25.2771    0.8209 0.3864E-01  1600.00  647.00  1437.997  10.965
   20    5.7506    0.1868   30.5485    0.9921 0.3864E-01  1600.00  647.00  1736.820  15.790
 Bottom                     33.4238    1.0855
 Lower layer of time doubling:   3  4  6  8 10 12 14 20    7 doublings, the limit for N2=384

 \vspace{-3.mm} 
\begin{verbatim}
v232
        DIFFI(J)=QK(J)/QR(J)    ! layer diffusivity
        K=1
        IF (LOCAL) K=J
        BLAY(J)= FLAY * RLAY**(J-1) * DSQRT(DIFFI(K)*PERSEC/PIVAL) ! thickness
...

      QQ=CONVF                  ! input safety factor
      IF (QQ.LT.0.8) QQ=1.      ! avoid unstable binary time division
      SAFE2=2.*QQ               ! safety convergence factor
      IF (IC2 .LE. N1-2) THEN   ! check safety at first modified layer
        QQ=BLAY(IC2)**2/(SAFE2*DTIM* DIFFI(IC2)) ! extra safety factor
        IF (QQ .LT. 1.) THEN    ! must increase layer thickness
          DO J=IC2,N1           ! set layer  IC just stable for local diffusivity
            BLAY(J)= RLAY**(J-IC2) * SQRT(SAFE2*DTIM *DIFFI(J) )
          ENDDO
          KM=0                  ! allow no time doubling above this layer
        ELSE                    ! allow upper layer changes
          KM=LOG(QQ)/LOG(2.)    ! max # of doublings before  IC
        ENDIF
      ENDIF
      IF (IDB2.GE.2) WRITE(*,*)'SAFE2,actual+',SAFE2,QQ,KM1,KM,DTIM
C
      DO J=2,N1                 ! layer loop
..
        SCONVG(J)=BLAY(J)**2/(2.*DTIMI* DIFFI(J)) ! Safety factor
        IF (J.EQ.IC2) KM=KM1    ! remove constraint on upper layers
        IF (K.LT.KM .AND. J.GT.2 .AND. MOD(KKK,2).EQ.0 
     &       .AND. SCONVG(J).GT. SAFE2) THEN ! increase time step
          DTIMI=2.*DTIMI
          SCONVG(J)=SCONVG(J)/2.

v321  BLAY(J)= FLAY * RLAY**(J-2)...
 
v331==v342: 
L296+
          DIFFI(J)=KTT(J)/(DENN(J)*CTT(J)) ! layer diffusivity
          K=1
          IF (LOCAL) K=J
          BLAY(J)= FLAY * RLAY**(J-2) * DSQRT(DIFFI(K)*PERSEC/PIVAL) ! thickness
L366+
if using a lower material, and its first layer not stable, make it just stable by factor CONVF
        IF (QQ .LT. ARC3) THEN    ! must increase layer thickness
          DO J=JJH,N1           ! set layer JJH just stable for local diffusivity
            BLAY(J)= RLAY**(J-JJH) * DSQRT(CONVF*2.*DTIM *DIFFI(J))
L376+
      SCONVG(1)=0.
      DO J=2,N1                 ! LAYER LOOP 1: compute safety factor
        XCEN(J)= XCEN(J-1)+ (BLAY(J)+BLAY(J-1))/2. ! center depth
        SCONVG(J)=BLAY(J)**2/(2.*DTIM* DIFFI(J)) ! Safety factor
L385+  
Loop 2    Find lowest safety factor at each layer or below; temporary use FBI
      DO J=N1-1,2,-1            ! LAYER LOOP 2: minimum here or below
        IF (SCONVG(J).LT.QQ) then 
          QQ= SCONVG(J)
          K=J
        ENDIF
        FBI(J) = QQ !  minimum safety factor at this layer or lower
   if lowest  lt ARC3, quit
L404+
      DO J=2,N1                 ! LAYER LOOP 3: Assign time doubling
        IF (K.LT.KM .AND. J.GT.3 .AND. MOD(KKK,2).EQ.0 
     &    .AND. FBI(J).GT. QA*4.*CONVF) THEN ! double time step for this layer
          DTIMI=2.*DTIMI
          K=K+1                 ! have new binary interval
          N1K(K)=J-1            ! bottom layer of prior interval
          KKK=KKK/2          ! number of time steps for this and deeper layers
          QA=QA*4.              ! 4^K   update forecast factor
        ENDIF
...
        SCONVG(J)=SCONVG(J)/QA  ! safety after time doubling

Then: replace factor of 4. on QA above with 2.

\end{verbatim}

KM,JJH,ARC3,CONVF,QQ,DTIM=   9   7 0.80100 3.00000     9.903 0.57722E+02
LZONE,LALCON,JJH, IK1:4=  F  T     7     1     0     0     0
 37   43413.549   43413.549
 36   32826.880   32826.880
 35   24821.838   24821.838
 34   18768.875   18768.875
 33   14191.966   14191.966
 32   10731.165   10731.165
 31    8114.303    8114.303
 30    6135.579    6135.579   3067.789= Converg.factor

...
 11      30.293      30.293
 10      22.906      22.906
  9      17.320      17.320
  8      13.096      13.096
  7       9.903       9.903   4.951   first of lower material
  6       7.488       7.488   3.744
  5       5.662       5.662   5.662
  4       4.281       4.281
  3       3.237       3.237
  2       2.448       2.448    2.448
 Time-doubling fraction:  0.55405408    
 1.035E+06=Dens*Cp   3.733E-08=Diffu.      0.0325=Scale      200.00=Inertia
 Beginning at layer   7  At      0.0219 m.
 1.588E+06=Dens*Cp   1.745E-06=Diffu.      0.2220=Scale     2097.20=Inertia
 table ident to thin9
